Compiler and Runtime Support for Functions-as-a-Service (FaaS)

### 2021-12-09 paper/paper.tex
[paper]: start eval, fix up figure placement, fix content

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{lipsum}
\usepackage{mwe}
\usepackage[parfill]{parskip}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\begin{figure}
    \centerline{\pagefig{figs/sys.pdf}}
    \caption{System overview and compilation pipeline. }  
	\label{fig:sys}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/pins.png} 
        \caption{Example compiler instrumentation for profiler callbacks in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}
        and \texttt{free}}  
	    \label{fig:pins}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.77\textwidth]{figs/sizes.png} 
        \caption{Example allocations that the static analysis can target in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:sizes}
    \end{minipage}\hfill
\end{figure}

@@

The allocator primarly takes the static analyses (see \ref{sa})
as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
\textit{pre-allocated} because all the information needed to allocate is known at
compile-time. These \ALLS\ are placed into a \textit{compiler-directed memory pool} that is
allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
of the workload, $W$, begins. The size of the pool is the sum of all statically 
analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
replaced by load of a compiler-determined offset of the base address of the pool.
An \ALL\ can only be selected for this transformation if it is not flagged by 
the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
can deduce the number of invocations of its parent function or of the allocation
itself (if it belongs to a loop, for instance). Otherwise, the data
in memory or the pointer itself may be overwritten accidentally. The latter condition
can be alleviated with aggressive inlining and function versioning. An example of 
this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
benefits to this approach: 1) alleviating function call overheads, 2) enabling 
prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
it before $W$ even begins. This can boost performance and simplifies memory management
for FaaS instances.

@@

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.77\textwidth]{figs/pool.png} 
        \caption{Example allocations (top) transformed to fetch from the compiler-directed
        memory pool (bottom; here, wrapped into the function \texttt{pool\_fetch}). The 
        basic functionality of the pool is shown to the right. In context of simplified LLVM 
        IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:pool}
    \end{minipage}\hfill
\end{figure}

@@

Initialization for the compiler-directed memory pool and bump allocators with known block size at 
compile-time are injected by the compiler into a runtime function marked with the attribute 
\texttt{constructor}. Although initialization is theoretically most performant when done in parallel
to FaaS startup, this was not implemented in the system prototype.

@@

\section{Evaluation}

>>>>

\usepackage{amsmath}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{array}
\usepackage{titlesec}
\usepackage{lipsum}
\usepackage{mwe}
\usepackage[parfill]{parskip}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\begin{figure}[htp]
    \centerline{\pagefig{figs/sys.pdf}}
    \caption{System overview and compilation pipeline. }  
	\label{fig:sys}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/pins.png} 
        \caption{Example compiler instrumentation for profiler callbacks in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}
        and \texttt{free}}  
	    \label{fig:pins}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.70\textwidth]{figs/sizes.png} 
        \caption{Example allocations that the static analysis can target in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:sizes}
    \end{minipage}\hfill
\end{figure}

@@

The allocator primarly takes the static analyses (see \ref{sa})
as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. If
an allocation is not selected as a candidate for any transformations, it is left unchanged.
First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
\textit{pre-allocated} because all the information needed to allocate is known at
compile-time. These \ALLS\ are placed into a \textit{compiler-directed memory pool} that is
allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
of the workload, $W$, begins. The size of the pool is the sum of all statically 
analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
replaced by load of a compiler-determined offset of the base address of the pool.
An \ALL\ can only be selected for this transformation if it is not flagged by 
the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
can deduce the number of invocations of its parent function or of the allocation
itself (if it belongs to a loop, for instance). Otherwise, the data
in memory or the pointer itself may be overwritten accidentally. The latter condition
can be alleviated with aggressive inlining and function versioning. An example of 
this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
benefits to this approach: 1) alleviating function call overheads, 2) enabling 
prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
it before $W$ even begins. This can boost performance and simplifies memory management
for FaaS instances.

@@

\begin{figure} [htp]
        \centering
        \includegraphics[width=0.77\textwidth]{figs/pool.png} 
        \caption{Example allocations (top) transformed to fetch from the compiler-directed
        memory pool (bottom; here, wrapped into the function \texttt{pool\_fetch}). The 
        basic functionality of the pool is shown to the right. In context of simplified LLVM 
        IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:pool}
\end{figure}

@@

Initialization for the compiler-directed memory pool and bump allocators (with known block size at 
compile-time) are injected by the compiler into a runtime function marked with the \texttt{constructor}
attribute. Although initialization is theoretically most performant when done in parallel
to FaaS startup, this was not implemented in the prototype.

@@

\section{Evaluation}
To evaluate the prototype, four kernels (matrix multiplication, compressed-row-storage conversion,
breadth first traversal, and Huffman coding -- see Figure~\ref{fig:apps}) were used. All testing
was performed on a 28-Core Intel Xeon CPU E5-2690 v4 machine. All runs were single-threaded.

\begin{figure}
        \small
        \centering
        \begin{tabular}{ | m{5cm} | m{7cm} | } 
          \hline
            Matrix Multiplication & Naive algorithm ; [60 x 30] x [30 x 120] \\ 
          \hline
            Compressed-Row-Storage (CSR) & Converts a [500 x 400] matrix to CSR form \\ 
          \hline
            Breadth First Traversal & Traversal on graph with 1000 vertices and records the traversal ; graph represented with adjancency matrix  \\ 
          \hline
            Huffman Coding & Builds a Huffman Tree for a block of "Lorem Ipsum" text \\ 
          \hline
        \end{tabular}
        \caption{Kernels used for evaluation.}
        \label{fig:apps}
\end{figure}

We assess the compiler transforms for the custom allocator both statically and dynamically. Figure X
shows the number of eligible static \ALLS\ in the baseline IR for each kernel and the number of \ALLS\
that were transformed to invoke the allocator.

Next, we evaluate the performance of the transformed binaries against the baseline binaries (i.e.
without transformation, compiled straight from source code) for each kernel. Each kernel was 
run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure X shows the comparative
overhead ratio (in terms of cycles).

Finally, we break down the execution times for each major piece of the allocator runtime and
highlight worst-case observed times (also in cycles). Figure X shows this for each essential
method of the runtime.

\begin{figure}
        \small
        \centering
        \begin{tabular}{ | m{5cm} | m{9cm} | m{3cm} | } 
          \hline
            \texttt{Init} & Initialize the global allocator (that manages bump allocators) and compiler directed pool & 20418 \\ 
          \hline
            \texttt{AddAllocator} & Initialize a new bump allocator and add it to the global allocator & 27170 \\ 
          \hline
            \texttt{AllocateFromPool} & Fetching a pointer from the compiler-directed memory pool (wrapped as a function) & 52 \\ 
          \hline
            \texttt{AllocateFromBump} & Fetching a pointer from a bump allocator & 1372 \\ 
          \hline
            \texttt{AllocateFromBumpWithRuntimeInit} & Fetching a pointer from a bump allocator with allocator initialization, if necessary & 1372 \\ 
          \hline
        \end{tabular}
        \caption{Worst-case cycle times for essential methods in allocator runtime.}
        \label{fig:apps}
\end{figure}

Note that the profiler, which outputs the \DWS\ and raw data on working set size 
every allocation/deallocation in the workload, is functional but not shown here.
